Topological Equivariant Artist Model for Visualization Library Architecture

### 2021-12-03 paper/main.tex
changed restriction symbols + set builder notation

% *** MATH PACKAGES ***
%% with other math-related packages, you may want to disable it.
\usepackage{amsmath, amsthm, amsfonts,amssymb,eulervm,xspace, mathtools}
\usepackage{mathrsfs} % math script fonts
\usepackage{relsize} %bigger
\usepackage{bm}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{remark}
\newtheorem{example}{Example}[section]
% *** SPECIALIZED LIST PACKAGES ***
\usepackage{xcolor}
\usepackage{algorithmic}
\usepackage[utf8]{inputenc}
\usepackage{subfig} %ieee does not like subfigure
\usepackage{multicol}
\usepackage{tikz}
\usetikzlibrary{cd} % commutative diagrams
\newtheorem{prop}{Proposition} %math?
\usepackage[switch]{lineno}
\renewcommand{\linenumberfont}{\normalfont\bfseries\small\color{lightgray}}
\usepackage{minted}
\setminted[python]{fontsize=\scriptsize, 
                   linenos,
                   numbersep=8pt,
                   autogobble, 
                   frame=lines,
                   framesep=3mm} 
% *** ALIGNMENT PACKAGES ***
\usepackage{array}
\usepackage{tabulary}
% IEEEtran contains the IEEEeqnarray family of commands

@@

and the set of all sections over \dbase\ is denoted $\Gamma(\dbase, \dtotal)$ . Sections can also be defined locally, meaning over an open set $\openset \subseteq \dtotal$ and the set of all sections over an arbitrary open set is $\Gamma(\openset, \dtotal)$. 
\begin{equation}
  \label{eq:related-work:continuity:sections}
  \begin{split}
  \cgamma{\opensetc}{\dtotalc|_{\opensetc}} \coloneqq \{&\dsectionc^{0}, \cdots, \dsectionc^{i}, \cdots, \dsectionc^{n}\} \\
   &\dsectionc: \opensetc\rightarrow \dtotalc|_{\openset}
  \end{split}
\end{equation}
Butler et al proposed that \cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992} sections could be used to model data sets since they are maps from a point on the continuity to a point in the fiber, and therefore they encapsulate both. The set of sections is the space of all data sets with the same continuity and fields.

\subsection{Sheaves}
\label{sec:related-work:continuity:sheaves}
Sheaves can be thought of as an "algebraic data structure" \cite{ghristElementaryAppliedTopology2014} for data that lives over topological spaces. A presheaf $\sheafc$ is a map from the open set to the set of sections $\sheafc: \opensetc \rightarrow \cgamma{\opensetc}{ \dtotalc|_{\opensetc}}$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}. The presheaf preserves inclusion maps $\iota$ between open sets and the inclusion map $\iota^*$ between the sets of sections.

\begin{equation}
  \label{eq:related-work:continuity:presheaf}
  \begin{tikzcd}[column sep=small]
    {\cgamma{\opensetc_1}{\dtotalc|_{\opensetc_1}}} & {} & {\cgamma{\opensetc_2}{\dtotalc|_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook'] \\& &\\
    \opensetc_1 \arrow[uu, "\sheafc_{\dtotal}"] \arrow[rr, "\iota"', hook] & {} \arrow[uu, "\sheafc_{\dtotal}" description, dotted] & \opensetc_2 \arrow[uu, "\sheafc_\dtotal"']                    
    \end{tikzcd}
\end{equation}

This means that a smaller space $\openset_1$ is included in a larger space $\openset_2$ and a function $\dsection \in \Gamma(\openset, \dtotal|_{\openset})$ that is continuous over a larger space $\openset_2$ is continuous over a subspace $\openset_1 \subset\openset_2$. Sheaves are presheafs where sets of sections are defined over unions of open sets over a topology $\bigcup_{j \in I} \openset)_j \in \dbase$ \cite{spanier1989algebraic, nlab:sheaf}. Sheaves provide a mathematical abstraction of data, the space it lives over, and relationships between subsets.

@@

For example, as illustrated in \autoref{fig:artist:gbase}, the display space is a 2D screen and the data space is a 1D line \dbase. For that line to have any thickness, it must be rendered as a band. This means the region $(\dbasepoint, \alpha)|_{\alpha \in I} \subset \gbase$ maps to the corresponding point $\dbasepoint \in \dbase$.

\subsubsection{Data}
\note{clean up why sheaf}
As mentioned in \autoref{sec:related-work:continuity:sheaves}, sheaves are a powerful tool for keeping track of data that lies over topological spaces. Therefore, we model the type of data an artist expects as a sheaf $\sheafc({\dtotalc}): \mathcal{\dbasec}^{op} \textcolor{sheaf}{\rightarrow} \setc$.
\begin{equation}
  \label{eq:artist:morphism-of-sheaf-data}
  \begin{tikzcd}
    & \cgamma{\opensetc}{\dtotalc|_{\opensetc}} \in Ob(\setc) & \\
    {} & & {} \\
    & \opensetc \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \arrow[uu, "\sheafc(\dtotalc)"', shift left = 10, color=sheaf] &   
\end{tikzcd}
\end{equation}
The objects of \textbf{Set} in this sheaf are the sets of sections $\Gamma(\dbase, \dtotal)$ introduced in \autoref{eq:related-work:continuity:sections} and the morphisms are the inclusion maps $(\iota, \iota^*)$ introduced in \autoref{eq:related-work:continuity:presheaf}.

@@

\subsubsection{Graphic}
We model the space of all graphic generating functions over a specific graphic topology $\opensetg \subseteq \gbase$ and targeting a specific display as a set of sections 
\begin{equation}
  \label{eq:artist:graphic_sections}
    \begin{split}
    \cgamma{\opensetgc}{\gtotalc|_{\opensetgc}} \coloneqq \{&\gsectionc^{0}, \cdots, \gsectionc^{i}, \cdots, \gsectionc^{n}\} \\
     &\gsectionc: \opensetgc\rightarrow \gtotalc|_{\opensetgc}
    \end{split}
  \end{equation}
\end{equation}
The set of sets of sections $\{\Gamma(\opensetg, \gtotal)\}|_{\opensetg \in gbase}$ is subcategory of \textbf{Set}; therefore the full space of graphics is a sheaf $\sheaf(\gtotal)$
\begin{equation}
  \label{eq:artist:artist_sheaf}
  \begin{tikzcd}[row sep=huge]
    {\cgamma{\opensetgc}{\gtotalc|_{\opensetgc}}}&  & 
    \textcolor{set}{\{}\gsectionc:\opensetgc\rightarrow \gtotalc|_{\opensetgc}\textcolor{set}{\}}
    \arrow[ll, "\iota"', hook', color=set] \\ &  & \\
    \opensetgc \arrow[uu, "\sheafc(\gtotalc)", color=sheaf] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheafc_{\vartistc}(\gtotalc)" description, dashed, color=sheaf] &  & 
    \end{tikzcd}
\end{equation}

\begin{equation}
  \label{eq:artist:artist_sheaf}
  \begin{tikzcd}[column sep=tiny, row sep=huge]
    {\cgamma{\opensetgc}{\gtotalc|_{\opensetgc}}}&  & 
    \textcolor{set}{\{}\gsectionc:\opensetgc\rightarrow \gtotalc|_{\opensetgc}\textcolor{set}{\}}
    \arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\mathlarger{\supset}}}}" description, hook', color=white] \\ &  & \\
    \opensetgc \arrow[uu, "\sheafc(\gtotalc)", color=sheaf] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheafc_{\vartistc}(\gtotalc)" description, dashed, color=sheaf] &  & 
    \end{tikzcd}
\end{equation}

The set of graphics reachable through an artist function is a subset of sections of the full set of graphics $\{\gsection\} \subset \Gamma(\opensetg, \gtotal|_{\opensetg})$. These sets of reachable graphics over open sets \opensetg\ are also objects of \textbf{Set}. This allows us to formulate the graphic output by the artist as a sheaf 
\begin{equation}
  \label{eq:artist:artist_sheaf}
  \sheafc_{\vartist}(\gtotalc): \gbasec^{\textcolor{base}{op}} \textcolor{sheaf}{\rightarrow} \setc
\end{equation}
from the graphic base space to the set of reachable graphics.

@@

\begin{definition} 
  \label{def:category:F} The fiber category $\mathcal{\dfiberc}$ is a monoidal category\cite{milewskiCategoryTheoryProgrammers}, which means it is a category equipped with a bifunctur $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$
  \begin{itemize}
    \item \textit{object} a monoid object \dfiber. A monoid is a set with a binary operator that is associative, closed, and for which the set contains an identity element $i$ \cite{nlan:monoid}.
    \item \textit{morphisms} $\dfiber \otimes \dfiber \rightarrow \dfiber$ and unit $i \rightarrow \dfiber$
  \end{itemize}
\end{definition}
The bifunctor $\otimes$ combines categories in a way that  preserves identity and composition \cite{fongInvitationAppliedCategory2019}. For example, given a pair of categories $\mathcal{\dfiber_a},\mathcal{\dfiber_b}$, the bifunctor yields a set of pairs $\{(\dfiber_a, \dfiber_b) | \dfiber_a \in \mathcal{\dfiber_a}, \dfiber_b \in \mathcal{\dfiber_b}\}$. For the associated morphisms $g: \mathcal{\dfiber_a}\rightarrow \mathcal{\dfiber_a}$ and $h: \mathcal{\dfiber_b}\rightarrow \mathcal{\dfiber_b}$, the function $(g\times h): (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b}) \rightarrow (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b})$ is pointwise $(g\times h)(\dfiber_a, \dfiber_b)\coloneqq (g(\dfiber_a), h(\dfiber_b))$. Because of the bifunctor, objects of the category $\mathcal{\dfiber}$ can be the product of any arbitrary number and types of categories, including \textbf{Set}, \textbf{Graph}, and topological spaces \textbf{Top}.

@@

When a fiber bundle is trivial, we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{\dbasepoint}, \dfiber_{\dbasepoint})$ over a point \dbasepoint\ as the basis of our equivariance. This is because a sheaf $\sheaf(\dtotal)$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk $\sheaf(\dtotal)|_{\dbasepoint}\coloneqq \lim\limits_{\openset\ni \dbasepoint} \Gamma(\openset, \dtotal)$ \cite{StalkSheaf2019}. The fiber space over a point is embedded in the stalk over that point $\dfiber_{\dbasepoint}\hookrightarrow \mathscr{\dfiber}_{\dbasepoint}$. The sheaf maps \dfunc\ restricted to the stalk are members of $hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$. The maps $\dfunc \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ include the types of data transformations described in \autoref{sec:related-work:equivariance}In \autoref{sec:artist:construction}, we assume that the fiber bundles are trivial.

@@

where the projection functions $i_{\dbase_a},i_{\dbase_b}$ behave such that $i_{\dbase_a}(\dbasepoint)|_{\dbasepoint \in \dbase_b} = \left[\dbasepoint\right]$, $i_{\dbase_b}(\dbasepoint)|_{\dbasepoint \in \dbase_b} = \left[\dbasepoint\right]$, meaning the projection functions yield equivalent points in the disjoint union $\dbase_a \bigsqcup_{\dbase_c} \dbase_b$. Data that is completely disjoint in the index, as shown in \autoref{fig:artist:union:disjoint_k}, is a special case where $\dbase_c=\varnothing$ and therefore there is no constraint on the artist with regards to how this data is rendered.

>>>>

% *** MATH PACKAGES ***
%% with other math-related packages, you may want to disable it.
\usepackage{amsmath, amsthm, amsfonts,amssymb,eulervm,xspace, mathtools}
\renewcommand{\restriction}{\mathord{\upharpoonright}} %restriction w/p space
\usepackage{mathrsfs} % math script fonts
\usepackage{relsize} %bigger
\usepackage{bm}
\theoremstyle{definition}
\newtheorem{definition}{Definition}[section]
\theoremstyle{remark}
\newtheorem{example}{Example}[section]
% *** SPECIALIZED LIST PACKAGES ***
\usepackage{xcolor}
\usepackage{algorithmic}
\usepackage[utf8]{inputenc}
\usepackage{subfig} %ieee does not like subfigure
\usepackage{multicol}
\usepackage{tikz}
\usetikzlibrary{cd} % commutative diagrams
\newtheorem{prop}{Proposition} %math?
\usepackage[switch]{lineno}
\renewcommand{\linenumberfont}{\normalfont\bfseries\small\color{lightgray}}
\usepackage{minted}
\setminted[python]{fontsize=\scriptsize, 
                   linenos,
                   numbersep=8pt,
                   autogobble, 
                   frame=lines,
                   framesep=3mm} 
% *** ALIGNMENT PACKAGES ***
\usepackage{array}
\usepackage{tabulary}
% IEEEtran contains the IEEEeqnarray family of commands

@@

and the set of all sections over \dbase\ is denoted $\Gamma(\dbase, \dtotal)$ . Sections can also be defined locally, meaning over an open set $\openset \subseteq \dtotal$ and the set of all sections over an arbitrary open set is $\Gamma(\openset, \dtotal)$. 
\begin{equation}
  \label{eq:related-work:continuity:sections}
  \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \coloneqq \big\{\dsectionc: \opensetc\rightarrow \dtotalc\restriction_{\openset} \; \bigm{\vert} \pi(\dsectionc(\dbasepointc)) = \dbasepointc\;\forall \dbasepointc \in \opensetc \big\}
\end{equation}
Butler et al proposed that \cite{butlerVisualizationModelBased1989,butlerVectorBundleClassesForm1992} sections could be used to model data sets since they are maps from a point on the continuity to a point in the fiber, and therefore they encapsulate both. The set of sections is the space of all data sets with the same continuity and fields.

\subsection{Sheaves}
\label{sec:related-work:continuity:sheaves}
Sheaves can be thought of as an "algebraic data structure" \cite{ghristElementaryAppliedTopology2014} for data that lives over topological spaces. A presheaf $\sheafc$ is a map from the open set to the set of sections $\sheafc: \opensetc \rightarrow \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}}$ \cite{spanier1989algebraic,SheafMathematics2021,nlab:presheaf}. The presheaf preserves inclusion maps $\iota$ between open sets and the inclusion map $\iota^*$ between the sets of sections.

\begin{equation}
  \label{eq:related-work:continuity:presheaf}
  \begin{tikzcd}[column sep=small]
    {\cgamma{\opensetc_1}{\dtotalc\restriction_{{\opensetc_1}}}} & {} & {\cgamma{\opensetc_2}{\dtotalc\restriction_{\opensetc_2}}} \arrow[ll, "\iota^*"', hook'] \\& &\\
    \opensetc_1 \arrow[uu, "\sheafc_{\dtotal}"] \arrow[rr, "\iota"', hook] & {} \arrow[uu, "\sheafc_{\dtotal}" description, dotted] & \opensetc_2 \arrow[uu, "\sheafc_\dtotal"']                    
    \end{tikzcd}
\end{equation}

This means that a smaller space $\openset_1$ is included in a larger space $\openset_2$ and a function $\dsection \in \Gamma(\openset, \dtotal\restriction_{\openset})$ that is continuous over a larger space $\openset_2$ is continuous over a subspace $\openset_1 \subset\openset_2$. Sheaves are presheafs where sets of sections are defined over unions of open sets over a topology $\bigcup_{j \in I} \openset)_j \in \dbase$ \cite{spanier1989algebraic, nlab:sheaf}. Sheaves provide a mathematical abstraction of data, the space it lives over, and relationships between subsets.

@@

For example, as illustrated in \autoref{fig:artist:gbase}, the display space is a 2D screen and the data space is a 1D line \dbase. For that line to have any thickness, it must be rendered as a band. This means the region $(\dbasepoint, \alpha)\vert_{\alpha \in I} \subset \gbase$ maps to the corresponding point $\dbasepoint \in \dbase$.

\subsubsection{Data}
\note{clean up why sheaf}
As mentioned in \autoref{sec:related-work:continuity:sheaves}, sheaves are a powerful tool for keeping track of data that lies over topological spaces. Therefore, we model the type of data an artist expects as a sheaf $\sheafc({\dtotalc}): \mathcal{\dbasec}^{op} \textcolor{sheaf}{\rightarrow} \setc$.
\begin{equation}
  \label{eq:artist:morphism-of-sheaf-data}
  \begin{tikzcd}
    & \cgamma{\opensetc}{\dtotalc\restriction_{\opensetc}} \in Ob(\setc) & \\
    {} & & {} \\
    & \opensetc \in Ob(\mathcal{\dbasec}^{\textcolor{base}{op}}) \arrow[uu, "\sheafc(\dtotalc)"', shift left = 10, color=sheaf] &   
\end{tikzcd}
\end{equation}
The objects of \textbf{Set} in this sheaf are the sets of sections $\Gamma(\dbase, \dtotal)$ introduced in \autoref{eq:related-work:continuity:sections} and the morphisms are the inclusion maps $(\iota, \iota^*)$ introduced in \autoref{eq:related-work:continuity:presheaf}.

@@

\subsubsection{Graphic}
We model the space of all graphic generating functions over a specific graphic topology $\opensetg \subseteq \gbase$ and targeting a specific display as a set of sections 
\begin{equation}
  \label{eq:artist:graphic_sections}}
    \cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}} \coloneqq \big\{\gsectionc: \opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} \bigm{\vert} \pi(\gsectionc(\gbasepointc)) = \gbasepointc \; \forall \gbasepointc \in \opensetgc \big\}
\end{equation}

The set of sets of sections $\{\Gamma(\opensetg, \gtotal)\}\vert_{{\opensetg \in gbase}}$ is subcategory of \textbf{Set}; therefore the full space of graphics is a sheaf $\sheaf(\gtotal)$

\begin{equation}
  \label{eq:artist:artist_sheaf}
  \begin{tikzcd}[column sep=tiny, row sep=huge]
    {\cgamma{\opensetgc}{\gtotalc\restriction_{\opensetgc}}}&  & 
    \textcolor{set}{\big\{}\gsectionc:\opensetgc\rightarrow \gtotalc\restriction_{\opensetgc} 
    \bigm{\vert} \vartist(\dsectionc) = \gsectionc \textcolor{set} \;, \dsectionc \in \sheafc(\dtotalc){\big\}} 
    \arrow[ll, "\textcolor{set}{\mathlarger{\mathlarger{\supset}}}" description, hook', color=white] \\ &  & \\
    \opensetgc \arrow[uu, "\sheafc(\gtotalc)", color=sheaf] \arrow[rruu, "\vartistc(\sheafc(\dtotalc))\coloneqq \sheafc_{\vartistc}(\gtotalc)" description, dashed, color=sheaf] &  & 
    \end{tikzcd}
\end{equation}

The set of graphics reachable through an artist function is a subset of sections of the full set of graphics $\{\gsection\} \subset \Gamma(\opensetg, \gtotal\restriction_{\opensetg})$. These sets of reachable graphics over open sets \opensetg\ are also objects of \textbf{Set}. This allows us to formulate the graphic output by the artist as a sheaf 
\begin{equation}
  \label{eq:artist:artist_sheaf}
  \sheafc_{\vartist}(\gtotalc): \gbasec^{\textcolor{base}{op}} \textcolor{sheaf}{\rightarrow} \setc
\end{equation}
from the graphic base space to the set of reachable graphics.

@@

\begin{definition} 
  \label{def:category:F} The fiber category $\mathcal{\dfiberc}$ is a monoidal category\cite{milewskiCategoryTheoryProgrammers}, which means it is a category equipped with a bifunctur $\otimes: \mathcal{\dfiber} \times \mathcal{\dfiber} \rightarrow \mathcal{\dfiber}$
  \begin{itemize}
    \item \textit{object} a monoid object \dfiber. A monoid is a set with a binary operator that is associative, closed, and for which the set contains an identity element $i$ \cite{nlan:monoid}.
    \item \textit{morphisms} $\dfiber \otimes \dfiber \rightarrow \dfiber$ and unit $i \rightarrow \dfiber$
  \end{itemize}
\end{definition}
The bifunctor $\otimes$ combines categories in a way that  preserves identity and composition \cite{fongInvitationAppliedCategory2019}. For example, given a pair of categories $\mathcal{\dfiber_a},\mathcal{\dfiber_b}$, the bifunctor yields a set of pairs $\{(\dfiber_a, \dfiber_b) \big{\mvert} \dfiber_a \in \mathcal{\dfiber_a}, \dfiber_b \in \mathcal{\dfiber_b}\}$. For the associated morphisms $g: \mathcal{\dfiber_a}\rightarrow \mathcal{\dfiber_a}$ and $h: \mathcal{\dfiber_b}\rightarrow \mathcal{\dfiber_b}$, the function $(g\times h): (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b}) \rightarrow (\mathcal{\dfiber_a}\times \mathcal{\dfiber_b})$ is pointwise $(g\times h)(\dfiber_a, \dfiber_b)\coloneqq (g(\dfiber_a), h(\dfiber_b))$. Because of the bifunctor, objects of the category $\mathcal{\dfiber}$ can be the product of any arbitrary number and types of categories, including \textbf{Set}, \textbf{Graph}, and topological spaces \textbf{Top}.

@@

When a fiber bundle is trivial, we can dispense with localization and directly consider the fiber maps $Hom(\dfiber_{\dbasepoint}, \dfiber_{\dbasepoint})$ over a point \dbasepoint\ as the basis of our equivariance. This is because a sheaf $\sheaf(\dtotal)$ over a limit of open sets that contain a point $\dbasepoint \in \dbase$ is approximately the same as a sheaf over a point \dbasepoint\ and is called the stalk $\sheaf(\dtotal)\restriction_{\dbasepoint}\coloneqq \lim\limits_{\openset\ni \dbasepoint} \Gamma(\openset, \dtotal)$ \cite{StalkSheaf2019}. The fiber space over a point is embedded in the stalk over that point $\dfiber_{\dbasepoint}\hookrightarrow \mathscr{\dfiber}_{\dbasepoint}$. The sheaf maps \dfunc\ restricted to the stalk are members of $hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$. The maps $\dfunc \in Hom_{\mathcal{\dfiber}}(\dfiber,\dfiber)$ include the types of data transformations described in \autoref{sec:related-work:equivariance}In \autoref{sec:artist:construction}, we assume that the fiber bundles are trivial.

@@

where the projection functions $i_{\dbase_a},i_{\dbase_b}$ behave such that $i_{\dbase_a}(\dbasepoint)\vert_{\dbasepoint \in \dbase_b} = \left[\dbasepoint\right]$, $i_{\dbase_b}(\dbasepoint)\vert_{\dbasepoint \in \dbase_b} = \left[\dbasepoint\right]$, meaning the projection functions yield equivalent points in the disjoint union $\dbase_a \bigsqcup_{\dbase_c} \dbase_b$. Data that is completely disjoint in the index, as shown in \autoref{fig:artist:union:disjoint_k}, is a special case where $\dbase_c=\varnothing$ and therefore there is no constraint on the artist with regards to how this data is rendered.
